Thermoacoustic heat pumps

### 2023-04-26 paper.tex
added space and domestic applications

\subsubsection{Efficiency}
Various designs have been tested numerically or experimentally since Ceperly first wrote his paper in 1979. The most promising are the looped-based designs as discussed above but efficiencies are still poor and insufficient for commercial viability. There are also different aims of these devices. Some traveling wave TAHP have been designed to liquefy hydrogen where cooling power is important while some designs focus more on room temperature applications where efficiency and power density are important\cite{WangRoomTemperature}.
\newpara{}
Tijani et al.\cite{spoelstraHighTemperature} built a traveling wave TAHP by building on the design discussed above. They coupled the TAHP with a TAE using the same principle but with a reversed cycle. This makes this device thermally driven. They managed to get to 40\% of Carnot's performance working between a range of 80 degrees and 10 degrees Celcius and pumping 200 W. This was an encouraging result for further study in this domain.
\newpara{}
Luo et al.\ built a refrigerator using the design discussed above\cite{LuoRefrigerator}. They also did this by coupling a traveling wave TAHP with a TAE in the same system but in a different way than the previous example. This device only reached a COP of 0.216 but managed to pump 469 W.
\newpara{}
More recently, Wang et al.\ estimated numerically that a refrigerator could be built with a cooling power of 6.35 kW and a COP of 3. This was built on the principle of putting multiple regenerators in series and thus creating a multistage refrigerator. Real-world experimentation is required however to validate these results.
\newpara{}
An in-depth study on the efficiency of traveling wave TAHP by Ueda et al. showed that there are three parameters that mainly effect the efficiency of the regenerator: the regenerator's installation position, length, and flow-channel radius\cite{uedaOptimization}. They concluded that by optimizing these three parameters a COP of above 60\% of the Carnot COP could be achieved. Tartibu investigated the effect of different geometries and thus configurations of traveling wave TAHP and tabulated different findings of different researchers\cite{TARTIBU2019102}. His study shows that whilst traveling wave TAHP are not sufficiently efficient yet, they show great promise and have much higher potential than their standing wave counterparts.

\section{Applications}
This section will discuss real-world applications of TAHP.\@ There are not a lot of real-world applications outside of academia because of the infancy of this technology.
\subsection{JWT cryocooler\cite{Petach2014MidII,Moore_2017,ross2022conceptual}}
The James-Webb telescope (JWT) is a space telescope launched by NASA on 25 December 2021. The purpose of JWT was to conduct infrared astrology. Due to redshift, the light coming from very old and distant celestial bodies have a wavelength in the infrared range. The temperature of the mid-infrared camera (MIRI) aboard the JWT has to be as close to 0 kelvin as possible to minimize the infrared radiation of the telescope itself and thus minimize the noise.
\newpara{}
The heat shield of the JWT only manages to keep the temperature of the sensor at about 37 kelvin but the sensor has an operating temperature of 6.2 kelvin. NASA chose a thermoacoustic cooler because reliability, high cooling power, and energy density were key. The cryocooler is a standing wave TAHP operating on helium with three stacks. A tube of helium is connected to the cold heat exchangers of the TAHP and is cooled down to about 17 kelvin which is still not enough which is still not sufficient. At the end of the tube, there is a small hole that drops the pressure of the helium. Because of this Joule-Thompson effect, the helium cools down to 6.2 kelvin to finally cool the MIRI.
\newpara{}

\subsection{Blue Heart}
\subsection{Frans}
\section{Future challenges and limitations}

\section{Conclusion}
While efficiencies are not yet at the same levels as conventional cooling/heating applications, one should keep in mind that TAHP offer other important advantages. For example, they use inert gases which make them environment friendly. They are also very flexible in driver methods. They can be powered thermally and mechanically. They also have nearly no moving parts which make them very suitable for environments where reliability is key.
\newpage
\bibliography{uni}
\bibliographystyle{unsrt}
\end{document}

>>>>

\subsubsection{Efficiency}
Various designs have been tested numerically or experimentally since Ceperly first wrote his paper in 1979. The most promising are the looped-based designs as discussed above but efficiencies are still poor and insufficient for commercial viability. There are also different aims of these devices. Some traveling wave TAHP have been designed to liquefy hydrogen where cooling power is important while some designs focus more on room temperature applications where efficiency and power density are important\cite{WangRoomTemperature}.
\newpara{}
Tijani et al.\cite{spoelstraHighTemperature} built a traveling wave TAHP by building on the design discussed above. They coupled the TAHP with a TAE using the same principle but with a reversed cycle. This makes this device thermally driven. They managed to get to 40\% of Carnot's performance working between a range of 80 degrees and 10 degrees Celcius and pumping 200 W. This was an encouraging result for further study in this domain.
\newpara{}
Luo et al.\ built a refrigerator using the design discussed above\cite{LuoRefrigerator}. They also did this by coupling a traveling wave TAHP with a TAE in the same system but in a different way than the previous example. This device only reached a COP of 0.216 but managed to pump 469 W.
\newpara{}
More recently, Wang et al.\ estimated numerically that a refrigerator could be built with a cooling power of 6.35 kW and a COP of 3. This was built on the principle of putting multiple regenerators in series and thus creating a multistage refrigerator. Real-world experimentation is required however to validate these results.
\newpara{}
An in-depth study on the efficiency of traveling wave TAHP by Ueda et al.\ showed that three parameters mainly affect the efficiency of the regenerator: the regenerator's installation position, length, and flow-channel radius\cite{uedaOptimization}. They concluded that by optimizing these three parameters a COP of above 60\% of the Carnot COP could be achieved. Tartibu investigated the effect of different geometries and thus configurations of traveling wave TAHP and tabulated different findings of different researchers\cite{TARTIBU2019102}. His study shows that whilst traveling wave TAHP are not sufficiently efficient yet, they show great promise and have much higher potential than their standing wave counterparts.

\section{Applications}
This section will discuss real-world applications of TAHP.\@ There are not a lot of real-world applications outside of academia because of the infancy of this technology.
\subsection{Space applications}
Thermoacoustic applications are a good fit for space missions because of their reliability and energy density. In this section, I list two important applications.
\subsubsection{Space shuttle get-away-special (GAS)\cite{AdeffSpace,powerofsound}}
Two standing-wave TAHP were designed, tested, and certified by NASA for its launch of the GAS space shuttle in 1992. One component was devised to cool the urine and blood samples of astronauts the other component was designed to cool electronic components. The TAHP had the stereo configuration as discussed above and used a mixture of helium and xenon as the working gas. The devices achieved 16\% of Cartnot's COP which was impressive considering the novelty of the technology. Due to an abrupt stop of funding the project was never realized. Due to the promising technology, it was later picked up by the US Navy to cool radar electronics.
\subsubsection{JWT cryocooler\cite{Petach2014MidII,Moore_2017,ross2022conceptual}}
The James-Webb telescope (JWT) is a space telescope launched by NASA on 25 December 2021. The purpose of JWT was to conduct infrared astrology. Due to redshift, the light coming from very old and distant celestial bodies have a wavelength in the infrared range. The temperature of the mid-infrared camera (MIRI) aboard the JWT has to be as close to 0 kelvin as possible to minimize the infrared radiation of the telescope itself and thus minimize the noise.
\newpara{}
The heat shield of the JWT only manages to keep the temperature of the sensor at about 37 kelvin but the sensor has an operating temperature of 6.2 kelvin. NASA chose a thermoacoustic cooler because reliability, high cooling power, and energy density were key. The cryocooler is a standing wave TAHP operating on helium with three stacks. A tube of helium is connected to the cold heat exchangers of the TAHP and is cooled down to about 17 kelvin which is still not enough which is still not sufficient. At the end of the tube, there is a small hole that drops the pressure of the helium. Because of this Joule-Thompson effect, the helium cools down to 6.2 kelvin to finally cool the MIRI.\@ So far, this cooler has worked successfully and is expected to for the coming 18 years.

\subsection{Domestic applications}
Almost 80\% of domestic heating in Europe is still being done with fossil fuels according to the EU\cite{EU}. There are two startups in Europe seeking to improve these numbers by making heat pumps a more accessible and environment-friendly option.
\subsection{Blue Heart\cite{blueheart}}
Blue Heart is a Dutch spinoff of the Organisation for Applied Scientific Research (TNO). A Blue Heart device would replace the cold circuit of a heat pump. High energy density of TAHP reduces the space needed for heat pump installations. They developed a traveling wave TAHP with helium as the working fluid. The acoustic power is generated by two pistons opposite to each other moving at the same frequency. The vibrations get canceled making the TAHP more silent than their conventional compression-based counterparts. Blue Heart claims that its device works efficiently with every temperature input and output and has a larger operation envelope than compression-based heat pumps. They also claim to be compatible with every type of heat source and to be suitable for new and existing houses.
\newpara{}
A Blue Heart device would operate with source temperatures between -20 degrees to 50 degrees Celsius and with sink temperatures between 7 degrees and 80 degrees Celsius. It would also have a capacity of 6 kW. The company plans to make the device commercially available in 2024.
\subsection{Equium\cite{equium}}
Equium is a French startup that offers to replace the heat pump entirely in contrast to Blue Heart. The company has not revealed the inner workings of their device but it is expected to also be a traveling wave TAHP working with helium. The company promises an impressive maintenance-free lifetime of 30 years. These devices are also planned to be commercially available in 2024
\section{Future challenges and limitations}

\section{Conclusion}
While efficiencies are not yet at the same levels as conventional cooling/heating applications, one should keep in mind that TAHP offers other important advantages. For example, they use inert gases which make them environment friendly. They are also very flexible in driver methods. They can be powered thermally and mechanically. They also have nearly no moving parts which makes them very suitable for environments where reliability is key.
\newpage
\bibliography{uni}
\bibliographystyle{unsrt}
\end{document}
